Visualizations for the manuscript

### 2022-05-14 metanalysis_visualization/manuscript_Figures.Rmd
reformating bootstrap, adding bootstrap estimations for all perdiods and for disorders groups

```{r}
median_table1_all <- length_hospitalisation_values_all %>% 
  group_by(period) %>% 
  summarise(median_hospitalisation = mean(len_hospitalisation)) %>% 
  pivot_wider(names_from = period, values_from = median_hospitalisation) %>% 
  mutate(Category = "Length of stays", 
         name = "median") %>% 
  select(Category, name, before_pandemic, during_pandemic) %>% 
  rename(before_pandemic_count = before_pandemic, 
         during_pandemic_count = during_pandemic)
median_table1_all$before_pandemic_count <- as.character(median_table1_all$before_pandemic_count)
median_table1_all$during_pandemic_count <- as.character(median_table1_all$during_pandemic_count)

@@

@@

coefficients_df <- patient_count_psy_period_psy_all %>% 
  group_by(siteid) %>% 
  group_modify(coefficient_slope_calculation,
               slope_name = c("time", "periodduring_pandemic:time")
               ) %>% 
  rename(observed_slope_coefficient = estimate)
```

## Define bootstrap functions

@@

```{r}
bootstrap_pvalue_calculation <- function(df, slope_name, ...) {
  z <- unique(df[["observed_slope_coefficient"]]) / sd(df[[slope_name]])
  pvalue_coeff_interaction <- round(pnorm(abs(z), lower.tail = FALSE) * 2, 3)
  pvalue_coeff_interaction <- ifelse( pvalue_coeff_interaction >= 0.01, round( pvalue_coeff_interaction, 2), pvalue_coeff_interaction)
  list("pvalue" = pvalue_coeff_interaction,
       "observed_slope_coefficient" = round(unique(df[["observed_slope_coefficient"]]), 2)
  )
}
```

#### Calculate bootstrap

@@

bootstrapped_coefficients_all <- lapply(bootstrapped, `[[`, 1) %>% 
  bind_rows() %>% 
  rename(before_slope = time, 
         difference_slope = `periodduring_pandemic:time`) %>% 
  mutate(during_slope = before_slope + difference_slope) %>% 
  select(iteration, siteid, everything())

@@

### Primary analysis

```{r}
df_coefficients_boot_primary <- bootstrapped_coefficients_all %>% 
  left_join(filter(coefficients_df, term == "periodduring_pandemic:time"),
            by = "siteid")

names_site_id <- unique(df_coefficients_boot_primary$siteid)

plot_annotations_primary <- df_coefficients_boot_primary %>% 
  group_by(siteid) %>% 
  group_map( bootstrap_pvalue_calculation, slope_name = "difference_slope" ) %>% 
  set_names(names_site_id) %>% 
  lapply(FUN = function(vec) {
    paste0("Coeffs differences: ",
           vec$observed_slope_coefficient,
           " (pvalue: ",
           format.pval(vec$pvalue,
                       digits = 2,
                       eps = 10**-3, 
                       nsmall = 2),
           ")")
  }) %>% 
  unlist()

plot_annotations_df_primary <- data.frame(
  siteid = names(plot_annotations_primary), 
  label = plot_annotations_primary
) %>%
  left_join( country_map ) 
```

@@

## Secondary analysis : slopes before/during

```{r}
pvalues_boot_before <- bootstrapped_coefficients_all %>% 
  left_join(filter(coefficients_df, term == "time"), by = "siteid") %>% 
  group_by(siteid) %>% 
  group_map( bootstrap_pvalue_calculation, slope_name = "before_slope" ) %>% 
  setNames(unique(bootstrapped_coefficients_all$siteid)) %>% 
  bind_rows(.id = "siteid")

pvalues_boot_after <- bootstrapped_coefficients_all %>% 
  left_join(coefficients_df %>%
              pivot_wider(names_from = term,
                          values_from = observed_slope_coefficient) %>% 
              mutate(observed_slope_coefficient = time + `periodduring_pandemic:time`) %>% 
              select(siteid, observed_slope_coefficient),
            by = "siteid") %>% 
  group_by(siteid) %>% 
  group_map( bootstrap_pvalue_calculation, slope_name = "during_slope" ) %>% 
  setNames(unique(bootstrapped_coefficients_all$siteid)) %>% 
  bind_rows(.id = "siteid")

@@

```{r}
trends_siteid <- df_coefficients_boot_primary %>% 
  group_by(siteid) %>% 
  summarise(mean_boot = mean(difference_slope),
            observed = unique(observed_slope_coefficient),
            var_boot = var(difference_slope), 
            z = mean_boot / sqrt(var_boot), 
            pvalue = pnorm(z, lower.tail = FALSE)*2 )
```

@@

## Forest plot meta library

### By site
```{r}
m.siteid <- metagen(TE = observed,
                 seTE = sqrt( var_boot ),
                 data = trends_siteid,
                 studlab = plotsLabel,
                 sm = "MN",
                 fixed = FALSE,
                 random = TRUE,
                 method.tau = "DL")

@@

### By country
```{r}
m.country <- metagen(TE = pred,
                 seTE = se,
                 data = trends_country,
                 studlab = country,
                 sm = "MN",
                 fixed = FALSE,
                 random = TRUE,
                 method.tau = "DL")

@@

### Disorder group

@@

perc_disorder_group_total <- perc_disorder_group_complete %>%
  group_by( time_p, disorder_group ) %>%
  summarise( total_diagnoses = sum(count_icd, na.rm = TRUE),
             dg_cases = sum(count_dg),
             percentage = (dg_cases / total_diagnoses) * 100, 
             country = "ALL", 
             siteid = "ALL") %>% 
  mutate(period = ifelse(time_p < pandemic_start_date,
                         "before_pandemic", 
                         "during_pandemic"))

@@

#### Bootstrapping results by disorder groups

```{r }
disorders_to_select <- c(
  "Anxiety Disorders", 
  "Depressive Disorders", 
  "Feeding and Eating Disorders", 
  "Suicide or Self-Injury"
)

bootstrapped_disorders <- perc_disorder_group_total %>%
  filter(disorder_group %in% disorders_to_select) %>% 
  group_by(disorder_group) %>% 
  mutate(month_year = format(time_p, "%m"),
         time = seq_along(time_p))  %>% 
 group_map(~ bootstrapping_slopes(.x,
                                  group_name = .y,
                                  bootstrap_size = bootstrap_size,
                                  count = "dg_cases", 
                                  percentage = "percentage"))

bootstrapped_coefficients_disorder <- lapply(bootstrapped_disorders, `[[`, 1) %>%
  bind_rows() %>% 
  rename(before_slope = time, 
         difference_slope = `periodduring_pandemic:time`)

@@

quantile_bootstrapped_fitted_disorder <- bootstrapped_fitted_disorder %>%
  # group_by(time_p, period, plotsLabel, sex) %>%
  group_by(time_p, period,  disorder_group) %>%
  summarise("2.5%" = quantile(predicted, 0.025),
            "50%" = median(predicted, 0.5),
            "97.5%" = quantile(predicted, 0.975)) %>%
  pivot_longer(cols = c("2.5%", "50%", "97.5%"),
               names_to = "quantiles")
coefficients_df_disorder <- perc_disorder_group_total %>% 
  group_by(disorder_group) %>% 
  mutate(month_year = format(time_p, "%m"),
         time = seq_along(time_p))  %>% 
  group_modify(coefficient_slope_calculation, 
               slope_name = "periodduring_pandemic:time"
               ) %>% 
  rename(observed_slope_coefficient = estimate)

df_coefficients_boot_disorder <- bootstrapped_coefficients_disorder %>% 
  left_join(coefficients_df_disorder, by = c("disorder_group"))
```

```{r annotation boostrap disorders}
names_disorder <- unique(df_coefficients_boot_disorder$disorder_group)

plot_annotations_disorder <- df_coefficients_boot_disorder %>% 
  group_by(disorder_group) %>% 
  group_map( bootstrap_pvalue_calculation, slope_name = "difference_slope" ) %>%
  set_names(names_disorder) %>% 
  lapply(FUN = function(vec) {
    paste0("Coeffs differences: ",
           vec$observed_slope_coefficient,
           " (pvalue: ",
           format.pval(vec$pvalue, 
                       eps = 10**-3),
           ")")
  }) %>% 
  unlist()
plot_annotations_disorder_df <- data.frame(
  disorder_group = names(plot_annotations_disorder), 
  label = plot_annotations_disorder
)
```

```{r}
# plot_annotations_df <- plot_annotations_df %>%
#   left_join( country_map )

bootstrapped_fitted_disorder %>% 
  mutate(period = recode_factor(period, "before_pandemic" = "Before pandemic onset", "during_pandemic" = "Since pandemic onset")) %>% 
  ggplot(aes(x = time_p, y = predicted, group = interaction(iteration, period))) +
  facet_wrap("disorder_group",
             scales = "free_y",
             labeller = label_wrap_gen(width = 35,multi_line = TRUE)) +
  geom_line(aes(color = period),
            alpha = 0.03) +
  geom_line(data = quantile_bootstrapped_fitted_disorder,
            aes(y = value, group = interaction(quantiles, period)),
            linetype = "dashed") +
  geom_line(data = quantile_bootstrapped_fitted_disorder[quantile_bootstrapped_fitted_disorder$quantiles == "50%", ],
            aes(y = value, group = period)) +
  geom_point(data = filter(perc_disorder_group_total,
                           disorder_group %in% disorders_to_select),
             aes(y = percentage, group=NULL),
             alpha = 0.5) +
  scale_colour_manual(values = cbPalette) +
  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
  guides(color = guide_legend(override.aes = list(alpha = 1))) +
  geom_label(data = plot_annotations_disorder_df, 
             aes(label=label), 
            x = -Inf,
            y = +Inf,
            hjust = 0,
            vjust = 1,
            size = 4,
            inherit.aes = FALSE) +
  labs(
    #title = "Interrupted Time Series Analysis: Percentage of patients with psychiatric conditions",
    # subtitle = "Parametric bootstrap estimation: (10000 iterations)",
       x = "Calendar dates",
       y = "Percentage",
       color = "Period") +
  theme(strip.text = element_text(size = 10),
        axis.text.x=element_text(angle = 45, hjust=1, size = 8),
        axis.text.y = element_text(size = 8), 
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        title = element_text(size = 14),
        plot.subtitle = element_text(size = 8, face = "italic"), 
        legend.position = "bottom") 
ggsave("figure3_boostrapped_disorders.png", height = 15, width = 25, units = "cm")

>>>>

```{r}
median_table1_all <- length_hospitalisation_values_all %>% 
  group_by(period) %>% 
  summarise(median_hospitalisation = median(len_hospitalisation)) %>% 
  pivot_wider(names_from = period, values_from = median_hospitalisation) %>% 
  mutate(Category = "Length of stays", 
         name = "median") %>% 
  select(Category, name, before_pandemic, during_pandemic) %>% 
  rename(before_pandemic_count = before_pandemic, 
         during_pandemic_count = during_pandemic)
median_table1_all$before_pandemic_count <- as.character(median_table1_all$before_pandemic_count)
median_table1_all$during_pandemic_count <- as.character(median_table1_all$during_pandemic_count)

@@

### Calculate LR coefficients

@@

observed_coefficients_df <- patient_count_psy_period_psy_all %>% 
  group_by(siteid) %>% 
  group_modify(coefficient_slope_calculation,
               slope_name = c("time", "periodduring_pandemic:time")
               ) %>% 
  rename(observed_slope_coefficient = estimate) %>% 
  pivot_wider(names_from = term,
              values_from = observed_slope_coefficient) %>% 
  rename(before_slope = time, 
         difference_slope = `periodduring_pandemic:time`) %>% 
  mutate(during_slope = before_slope + difference_slope) %>% 
  pivot_longer(cols = c(before_slope, difference_slope, during_slope), 
               names_to = "term", 
               values_to = "observed_slope_coefficient")
```

### Define bootstrap functions

@@

```{r}
bootstrap_pvalue_calculation <- function(df, ...) {
  z <- unique(df[["observed_slope_coefficient"]]) / sd(df[["bootstrapped_slope_coefficient"]])
  pvalue_coeff_interaction <- round(pnorm(abs(z), lower.tail = FALSE) * 2, 3)
  pvalue_coeff_interaction <- ifelse( pvalue_coeff_interaction >= 0.01, round( pvalue_coeff_interaction, 2), pvalue_coeff_interaction)
  data.frame("pvalue" = pvalue_coeff_interaction,
       "observed_slope_coefficient" = round(unique(df[["observed_slope_coefficient"]]), 2)
  )
}
```

#### Calculate bootstrap: primary analysis

@@

bootstrapped_coefficients_all <- lapply(bootstrapped, `[[`, 1) %>% 
  bind_rows() %>% 
  rename(before_slope = time, 
         difference_slope = `periodduring_pandemic:time`) %>% 
  mutate(during_slope = before_slope + difference_slope) %>% 
  select(iteration, siteid, everything()) %>% 
    pivot_longer(cols = c(before_slope, difference_slope, during_slope), 
               names_to = "term", 
               values_to = "bootstrapped_slope_coefficient") %>% 
  left_join(observed_coefficients_df, by = c("siteid", "term"))

pvalues_boot <- bootstrapped_coefficients_all %>% 
  group_by(siteid, term) %>% 
  group_modify( bootstrap_pvalue_calculation, .keep = TRUE) %>% 
  mutate(label = paste0("Coeffs differences: ",
                        observed_slope_coefficient,
                        " (pvalue: ",
                        format.pval(pvalue, digits = 2, eps = 10**-2), ")")
  )

ci_boot <- bootstrapped_coefficients_all %>% 
  group_by(siteid, term) %>% 
  summarise("2.5%" = quantile(bootstrapped_slope_coefficient, 0.025), 
            "50%" = quantile(bootstrapped_slope_coefficient, 0.50), 
            "97.5%" = quantile(bootstrapped_slope_coefficient, 0.975)) %>% 
  mutate(ci = paste0("[", as.character(round(`2.5%`, 2)), "; ", as.character(round(`97.5%`, 2)), "]"))

@@

#### Trend differences

```{r}

plot_annotations_df_primary <- pvalues_boot %>%
  select(siteid, label) %>% 
  filter(term == "difference_slope") %>% 
  left_join(country_map)
```

@@

## Table with bootstrapped coefficients, and pvalues

```{r}
pvalues_boot %>% 
  inner_join(ci_boot) %>% 
  inner_join(country_map) %>% 
  arrange(plotsLabel) %>% 
  mutate(string = paste0(observed_slope_coefficient, " ", ci, " (", format.pval(pvalue, digits = 3, eps = 10**-2), ")")) %>% 
  select(plotsLabel, string) %>% 
  pivot_wider(names_from = term, values_from = string) %>% 
  select(plotsLabel, before_slope, during_slope, difference_slope) %>% 
  rename("Before pandemic trend [95% CI] (p-value)" = before_slope,
        "Since pandemic trend [95% CI] (p-value)" = during_slope,
        "Difference before/since pandemic trend [95% CI] (p-value)" = difference_slope)

@@

```{r}
trends_siteid <- bootstrapped_coefficients_all %>% 
  filter(term == "difference_slope") %>% 
  group_by(siteid) %>% 
  summarise(mean_boot = mean(bootstrapped_slope_coefficient),
            observed = unique(observed_slope_coefficient),
            var_boot = var(bootstrapped_slope_coefficient), 
            z = mean_boot / sqrt(var_boot), 
            pvalue = pnorm(z, lower.tail = FALSE)*2 )
```

@@

### Meta analysis with another package

#### Site
```{r}
m.siteid <- metagen(TE = observed,
                 seTE = sqrt( var_boot ),
                 data = trends_siteid,
                 studlab = plotsLabel,
                 sm = "MN",
                 fixed = FALSE,
                 random = TRUE,
                 method.tau = "DL")

@@

#### By country
```{r}
m.country <- metagen(TE = pred,
                 seTE = se,
                 data = trends_country,
                 studlab = country,
                 sm = "MN",
                 fixed = FALSE,
                 random = TRUE,
                 method.tau = "DL")

@@

### Secondary analysis: :disorder group

@@

disorders_to_select <- c(
  "Anxiety Disorders", 
  "Depressive Disorders", 
  "Feeding and Eating Disorders", 
  "Suicide or Self-Injury"
)

perc_disorder_group_total <- perc_disorder_group_complete %>%
  group_by( time_p, disorder_group ) %>%
  summarise( total_diagnoses = sum(count_icd, na.rm = TRUE),
             dg_cases = sum(count_dg),
             percentage = (dg_cases / total_diagnoses) * 100, 
             country = "ALL", 
             siteid = "ALL") %>% 
  mutate(period = ifelse(time_p < pandemic_start_date,
                         "before_pandemic", 
                         "during_pandemic")) %>% 
    filter(disorder_group %in% disorders_to_select)

@@

#### Observed results: disorders

```{r}
observed_coefficients_df_disorder <- perc_disorder_group_total %>% 
  group_by(disorder_group) %>% 
  mutate(month_year = format(time_p, "%m"),
         time = seq_along(time_p))  %>% 
  group_modify(coefficient_slope_calculation, 
               slope_name = c("time", "periodduring_pandemic:time")
               ) %>% 
  rename(observed_slope_coefficient = estimate) %>% 
  pivot_wider(names_from = term,
              values_from = observed_slope_coefficient) %>% 
  rename(before_slope = time, 
         difference_slope = `periodduring_pandemic:time`) %>% 
  mutate(during_slope = before_slope + difference_slope) %>% 
  pivot_longer(cols = c(before_slope, difference_slope, during_slope), 
               names_to = "term", 
               values_to = "observed_slope_coefficient")

#### Calculate bootstrap: disorders groups

```{r}
# Get count_icd_all in the same format that patient_count_psy_period_psy_all, because first n last months missing ...

bootstrapped_disorders <- perc_disorder_group_total %>%
  group_by(disorder_group) %>% 
  mutate(month_year = format(time_p, "%m"),
         time = seq_along(time_p))  %>% 
  group_map( ~bootstrapping_slopes(.x,
                                   group_name = .y,
                                   bootstrap_size = bootstrap_size,
                                   count = "dg_cases", 
                                   percentage = "percentage"))
names(bootstrapped_disorders) <- unique(perc_disorder_group_total$disorder_group)

bootstrapped_coefficients_disorders <- lapply(bootstrapped_disorders, `[[`, 1) %>% 
  bind_rows() %>% 
  rename(before_slope = time, 
         difference_slope = `periodduring_pandemic:time`) %>% 
  mutate(during_slope = before_slope + difference_slope) %>% 
  select(iteration, disorder_group, everything()) %>% 
    pivot_longer(cols = c(before_slope, difference_slope, during_slope), 
               names_to = "term", 
               values_to = "bootstrapped_slope_coefficient") %>% 
  left_join(observed_coefficients_df_disorder, by = c("disorder_group", "term"))

pvalues_boot_disorders <- bootstrapped_coefficients_disorders %>% 
  group_by(disorder_group, term) %>% 
  group_modify( bootstrap_pvalue_calculation, .keep = TRUE) %>% 
  mutate(label = paste0("Coeffs differences: ",
                        observed_slope_coefficient,
                        " (pvalue: ",
                        format.pval(pvalue, digits = 2, eps = 10**-2), ")")
  )

ci_boot_disorders <- bootstrapped_coefficients_disorders %>% 
  group_by(disorder_group, term) %>% 
  summarise("2.5%" = quantile(bootstrapped_slope_coefficient, 0.025), 
            "50%" = quantile(bootstrapped_slope_coefficient, 0.50), 
            "97.5%" = quantile(bootstrapped_slope_coefficient, 0.975)) %>% 
  mutate(ci = paste0("[", as.character(round(`2.5%`, 2)), "; ", as.character(round(`97.5%`, 2)), "]"))

@@

quantile_bootstrapped_fitted_disorder <- bootstrapped_fitted_disorder %>%
  group_by(time_p, period, disorder_group) %>%
  summarise("2.5%" = quantile(predicted, 0.025),
            "50%" = median(predicted, 0.5),
            "97.5%" = quantile(predicted, 0.975)) %>%
  pivot_longer(cols = c("2.5%", "50%", "97.5%"),
               names_to = "quantiles")

#### Table coefficients disorders

```{r}
pvalues_boot_disorders %>% 
  inner_join(ci_boot_disorders) %>% 
  arrange(disorder_group) %>% 
  mutate(string = paste0(observed_slope_coefficient, " ", ci, " (", format.pval(pvalue, digits = 3, eps = 10**-2), ")")) %>% 
  select(disorder_group, string) %>% 
  pivot_wider(names_from = term, values_from = string) %>% 
  select(disorder_group, before_slope, during_slope, difference_slope) %>% 
  rename("Before pandemic trend [95% CI] (p-value)" = before_slope,
        "Since pandemic trend [95% CI] (p-value)" = during_slope,
        "Difference before/since pandemic trend [95% CI] (p-value)" = difference_slope) %>% 
  flextable::flextable() %>% 
  flextable::save_as_docx(path = "table_disorders.docx")

```{r}

plot_annotations_disorder_df <-  pvalues_boot_disorders %>%
  select(disorder_group, label) %>% 
  filter(term == "difference_slope")
```

```{r}
bootstrapped_fitted_disorder %>% 
  mutate(period = recode_factor(period, "before_pandemic" = "Before pandemic onset", "during_pandemic" = "Since pandemic onset")) %>% 
  ggplot(aes(x = time_p, y = predicted, group = interaction(iteration, period))) +
  facet_wrap("disorder_group",
             scales = "free_y",
             labeller = label_wrap_gen(width = 35,multi_line = TRUE)) +
  geom_line(aes(color = period),
            alpha = 0.03) +
  geom_line(data = quantile_bootstrapped_fitted_disorder,
            aes(y = value, group = interaction(quantiles, period)),
            linetype = "dashed") +
  geom_line(data = quantile_bootstrapped_fitted_disorder[quantile_bootstrapped_fitted_disorder$quantiles == "50%", ],
            aes(y = value, group = period)) +
  geom_point(data = filter(perc_disorder_group_total,
                           disorder_group %in% disorders_to_select),
             aes(y = percentage, group=NULL),
             alpha = 0.5) +
  scale_colour_manual(values = cbPalette) +
  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
  guides(color = guide_legend(override.aes = list(alpha = 1))) +
  geom_label(data = plot_annotations_disorder_df, 
             aes(label=label), 
            x = -Inf,
            y = +Inf,
            hjust = 0,
            vjust = 1,
            size = 4,
            inherit.aes = FALSE) +
  labs(
    #title = "Interrupted Time Series Analysis: Percentage of patients with psychiatric conditions",
    # subtitle = "Parametric bootstrap estimation: (10000 iterations)",
       x = "Calendar dates",
       y = "Percentage",
       color = "Period") +
  theme(strip.text = element_text(size = 10),
        axis.text.x=element_text(angle = 45, hjust=1, size = 8),
        axis.text.y = element_text(size = 8), 
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        title = element_text(size = 14),
        plot.subtitle = element_text(size = 8, face = "italic"), 
        legend.position = "bottom") 
ggsave("figure3_boostrapped_disorders.png", height = 15, width = 25, units = "cm")
